### 2016-12-25 Letters/Letter1.tex

\documentclass[11pt]{article}

%\usepackage{fullpage}
\usepackage[top=5cm, bottom=1cm, left=2.5cm, right=2.5cm]{geometry}

\begin{document}
\begin{Large}

\begin{flushleft}
The Manager,\\
State Bank of India,\\
Serampore Branch,\\
Serampore,\\
Hooghly-712201.\\
\bigbreak
Dear Sir,\\
\end{flushleft}
\begin{center}
\underline{Sub:Request to marry your daughter.}
\end{center}

%\justify
This is to inform you that, I, Rohit Das, of 108, Netaji Subhas Avenue, Serampore, Hooghly-712201, would like to marry your daughter.\\
\newline
\ \ \ We have been in a romantic relationship for the last 5 years, and now that I am earning for myself, I would like to take the responsibility of happiness and well-being of your elder daughter, Ms. Sukanya Saha. I am earning a decent amount of Rs. 7lpa at a reputed company, the likes of which your daughter can describe to you.
\ \ \

\end{Large} 
\end{document}