### 2017-08-03 cv/cv-sections/education.tex
Adding CV
Signed-off-by: mouri11 <rohit.das950@gmail.com>

%----------------------------------------------------------------------------------------
%	SECTION TITLE
%----------------------------------------------------------------------------------------

\cvsection{Education}

%----------------------------------------------------------------------------------------
%	SECTION CONTENT
%----------------------------------------------------------------------------------------

\begin{cventries}

%------------------------------------------------

\cventry
{B. Tech.  in Computer Science and Engineering} % Degree
{MAKAUT, WB (Maulana Abul Kalam Azad University of Technology)\newline (formerly WBUT)} % Institution
{BF -142, Salt Lake Sector -I, \newline
Kolkata - 700064.} % Location
{Jul. 2014 - PRESENT} % Date(s)
{ % Description(s) bullet points
\begin{cvitems}
\item {Got cleared under TFW scheme (Tuition-Fee Waiver scheme) meant for students ranking well in WBJEE.}
\end{cvitems}
}

%------------------------------------------------

\cventry
{Higher Secondary Education} % Degree
{Holy Home School} % Institution
{Serampore - 712201.} % Location
{Apr. 2012 - Mar. 2014} % Date(s)
{ % Description(s) bullet points
\begin{cvitems}
\item {Indian School Certificate Examination (ISC) score - 95.25\%.}
\end{cvitems}
}

%------------------------------------------------

\cventry
{Secondary Education} % Degree
{Don Bosco School} % Institution
{Bandel, P. O. \& Dist. Hooghly - 712103.} % Location
{Apr. 2002 - Mar. 2012} % Date(s)
{ % Description(s) bullet points
\begin{cvitems}
\item {Indian Certificate of Secondary Education (ICSE) score - 94.4\%.}
\item {Qualified for Anant Scholarship, provided to meritorious students passing class 10 or equivalent.}
\end{cvitems}
}

%------------------------------------------------

\end{cventries}