### 2020-01-20 cv/cv-sections/education.tex
cv.pdf: Added MTech Creds
Added marks for MTech at IIT Bhilai, and details of my project on
Edge computing.

%----------------------------------------------------------------------------------------
%	SECTION TITLE
%----------------------------------------------------------------------------------------

\cvsection{Education}

%----------------------------------------------------------------------------------------
%	SECTION CONTENT
%----------------------------------------------------------------------------------------

\begin{cventries}

%------------------------------------------------

\cventry
{M. Tech.  in Computer Science and Engineering} % Degree
{Indian Institute of Technology, Bhilai} % Institution
{GEC Campus, Old Dhamtari Rd, Sejbahar, Chhattisgarh - 492015} % Location
{Jul. 2019 - Present} % Date(s)
{ % Description(s) bullet points
\begin{cvitems}
\item {CGPA (2019-20 S) - 9.75}
\item {Currently working on Data Science using Edge Computing}
\item {Outreach (Placement and Training) Cell Volunteer}
\end{cvitems}
}

%------------------------------------------------

\cventry
{B. Tech.  in Computer Science and Engineering} % Degree
{MAKAUT, WB (Maulana Abul Kalam Azad University of Technology)\newline (formerly WBUT)} % Institution
{BF -142, Salt Lake Sector -I, \newline
Kolkata - 700064.} % Location
{Jul. 2014 - Jul. 2018} % Date(s)
{ % Description(s) bullet points
\begin{cvitems}
\item {DGPA - 8.11}
\item {Got cleared under TFW scheme (Tuition-Fee Waiver scheme) meant for students ranking well in WBJEE.}
\end{cvitems}
}

%------------------------------------------------

\cventry
{Higher Secondary Education} % Degree
{Holy Home School, Serampore} % Institution
{Serampore - 712201.} % Location
{Apr. 2012 - Mar. 2014} % Date(s)
{ % Description(s) bullet points
\begin{cvitems}
\item {Indian School Certificate Examination (ISC) score - 95.33\%.}
\item {Qualified for Anant Scholarship, provided to meritorious students passing class 10 or equivalent.}
\end{cvitems}
}

%------------------------------------------------

%\cventry
%{Secondary Education} % Degree
%{Don Bosco School, Bandel} % Institution
%{Bandel, P. O. \& Dist. Hooghly - 712103.} % Location
%{Apr. 2002 - Mar. 2012} % Date(s)
%{ % Description(s) bullet points
%\begin{cvitems}
%\item {Indian Certificate of Secondary Education (ICSE) score  - 93.71\%.}
%\end{cvitems}
%}

%------------------------------------------------

\end{cventries}